An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-02-22 EconGrowth.Rnw
Now including a cesPredictions function with example using US data.

<<ces functions, eval=TRUE>>=
####################
# Returns an nls ces model for the country and energyType specified.
# energyType should be one of Q", "X", or "U".
##
cesModel <- function(countryAbbrev, energyType){
  # Load the data that we need.
  dataTable <- loadData(countryAbbrev)
  if (energyType == "U"){
    # Trim the dataset to include only those years for which U is available.
    dataTable <- subset(dataTable, !is.na(iU))
  }
  # Trim data from China and South Africa that is missing labor information for 2011.
  dataTable <- subset(dataTable, !is.na(iLabor))
  # We need to do the CES fit with the desired energyType.
  # To achieve the correct fit, we'll change the name of the desired column
  # to "iEToFit" and use "iEToFit" in the nls function.
  dataTable <- replaceColName(dataTable, energyType, "iEToFit")
  # Establish default guesses for sigma, gamma_E, beta, lambda_L and lambda_E.
  start <- list(sigma=0.9, gamma_E=0.01, beta=0.7, lambda_L=0.011, lambda_E=0.018) #values for Sweden
  # Establish bounds
  lower = list(sigma=-Inf, gamma_E=0.0, beta=-Inf, lambda_L=-Inf, lambda_E=-Inf)
  upper = list(sigma= Inf, gamma_E=1.0, beta= Inf, lambda_L= Inf, lambda_E= Inf)
  # Now some country-specific values.
  if (countryAbbrev == "US"){
    start <- list(sigma=0.95, gamma_E=0.1, beta=0.7, lambda_L=0.015, lambda_E=0.015) #values for Sweden
  }
  if (countryAbbrev == "UK"){
    start <- list(sigma=0.95, gamma_E=0.2, beta=0.7, lambda_L=0.02, lambda_E=0.02) #values for Sweden
    lower = list(sigma=-Inf, gamma_E=0.0, beta=-Inf, lambda_L=-0.3, lambda_E=-0.3)
    upper = list(sigma= Inf, gamma_E=1.0, beta= Inf, lambda_L= 0.3, lambda_E= 0.3)
  }
  if (countryAbbrev == "CN"){

  }
  # Runs a non-linear least squares fit to the data with constraints
  # This is the version of the function from the Stern and Kander (2012) paper. However, it includes a 
  # 1/sigma exponent on the gamma_E terms, which, I believe, is bogus.
#   model <- iGDP ~ ((1-gamma_E)^(1.0/sigma) * ((exp(lambda_L*iYear))^beta 
#                                                   * iLabor^beta
#                                                   * iCapStk^(1-beta) )^((sigma-1.0)/sigma) 
#                   + gamma_E^(1.0/sigma)*(exp(lambda_E*iYear) * iEToFit)^((sigma-1.0)/sigma))^(sigma/(sigma-1.0))

  # This is the corrected version of the Stern and Kander (2012) paper.
#   model <- iGDP ~ ((1-gamma_E) * ((exp(lambda_L*iYear))^beta * iLabor^beta * iCapStk^(1-beta) )^((sigma-1.0)/sigma) 
#                   + gamma_E * (exp(lambda_E*iYear) * iEToFit)^((sigma-1.0)/sigma))^(sigma/(sigma-1.0))

  # This is the nls fit to the CES model  
#   modelCES <- nls(formula=model, data=dataTable, start=start,
#                   control=nls.control(maxiter=10000, tol=1e-05, minFactor=1/1024, 
#                                       printEval=TRUE, warnOnly=TRUE),
#                   algorithm = "port",
#                   lower = lower,
#                   upper = upper,
#                   trace=TRUE
#                   )

  # An attempt with nlmrt function nlxb. Doesn't provide many good results.
#   modelCES <- nlxb(model,
#                    data=dataTable,
#                    start=list(sigma=sigmaGuess, gamma_E=gamma_EGuess, beta=betaGuess, 
#                               lambda_L=lambda_LGuess, lambda_E=lambda_EGuess),
#                    trace=TRUE
#                   )

@@

@@

@@

  return(modelCES)
}
# print("US U")
# print(summary(cesModel("US", "U")))
# 
# print("")
# print("UK U")
# print(summary(cesModel("UK", "U")))
# 
# print("")
# print("JP U")
# print(summary(cesModel("JP", "U")))

# print("")
# print("CN Q")
# print(summary(cesModel("CN", "Q")))
# 
# print("")
# print("ZA Q")
# print(summary(cesModel("ZA", "Q")))
# 
# print("")
print("SA Q")
print(summary(cesModel("SA", "Q")))
# 
# print("")
# print("IR Q")
# print(summary(cesModel("IR", "Q")))
# 
# print("")
# print("TZ Q")
# print(summary(cesModel("TZ", "Q")))
# 
# print("")
# print("ZM Q")
# print(summary(cesModel("ZM", "Q")))

>>>>

<<ces functions, eval=TRUE>>=
######################################
# Loads data and trims rows where necessary.
# Also, replaces the name of the desired energy column with "iEToFit".
# Returns a data.frame with the loaded data.
##
loadAndPrepDataForCES <- function(countryAbbrev, energyType){
  dataTable <- loadData(countryAbbrev)
  if (energyType == "U"){
    # Trim the dataset to include only those years for which U is available.
    dataTable <- subset(dataTable, !is.na(iU))
  }
  # Trim data from China and South Africa that is missing labor information for 2011.
  dataTable <- subset(dataTable, !is.na(iLabor))
  # We need to do the CES fit with the desired energyType.
  # To achieve the correct fit, we'll change the name of the desired column
  # to "iEToFit" and use "iEToFit" in the nls function.
  dataTable <- replaceColName(dataTable, energyType, "iEToFit")
  return(dataTable)
}

####################
# Returns an nls ces model for the country and energyType specified.
# energyType should be one of Q", "X", or "U".
##
cesModel <- function(countryAbbrev, energyType){
  # Load the data that we need.
  dataTable <- loadAndPrepDataForCES(countryAbbrev, energyType)
#   # Establish default guesses for sigma, gamma_E, beta, lambda_L and lambda_E.
#   start <- list(sigma=0.9, gamma_E=0.01, beta=0.7, lambda_L=0.011, lambda_E=0.018) #values for Sweden
#   # Establish bounds
#   lower = list(sigma=-Inf, gamma_E=0.0, beta=-Inf, lambda_L=-Inf, lambda_E=-Inf)
#   upper = list(sigma= Inf, gamma_E=1.0, beta= Inf, lambda_L= Inf, lambda_E= Inf)
#   # Now some country-specific values.
#   if (countryAbbrev == "US"){
#     start <- list(sigma=0.95, gamma_E=0.1, beta=0.7, lambda_L=0.015, lambda_E=0.015) #values for Sweden
#   }
#   if (countryAbbrev == "UK"){
#     start <- list(sigma=0.95, gamma_E=0.2, beta=0.7, lambda_L=0.02, lambda_E=0.02) #values for Sweden
#     lower = list(sigma=-Inf, gamma_E=0.0, beta=-Inf, lambda_L=-0.3, lambda_E=-0.3)
#     upper = list(sigma= Inf, gamma_E=1.0, beta= Inf, lambda_L= 0.3, lambda_E= 0.3)
#   }
#   if (countryAbbrev == "CN"){
#     
#   }
#   # Runs a non-linear least squares fit to the data with constraints
#   # This is the version of the function from the Stern and Kander (2012) paper. However, it includes a 
#   # 1/sigma exponent on the gamma_E terms, which, I believe, is bogus.
# #   model <- iGDP ~ ((1-gamma_E)^(1.0/sigma) * ((exp(lambda_L*iYear))^beta 
# #                                                   * iLabor^beta
# #                                                   * iCapStk^(1-beta) )^((sigma-1.0)/sigma) 
# #                   + gamma_E^(1.0/sigma)*(exp(lambda_E*iYear) * iEToFit)^((sigma-1.0)/sigma))^(sigma/(sigma-1.0))
#   
#   # This is the corrected version of the Stern and Kander (2012) paper.
# #   model <- iGDP ~ ((1-gamma_E) * ((exp(lambda_L*iYear))^beta * iLabor^beta * iCapStk^(1-beta) )^((sigma-1.0)/sigma) 
# #                   + gamma_E * (exp(lambda_E*iYear) * iEToFit)^((sigma-1.0)/sigma))^(sigma/(sigma-1.0))
#   
#   # This is the nls fit to the CES model  
# #   modelCES <- nls(formula=model, data=dataTable, start=start,
# #                   control=nls.control(maxiter=10000, tol=1e-05, minFactor=1/1024, 
# #                                       printEval=TRUE, warnOnly=TRUE),
# #                   algorithm = "port",
# #                   lower = lower,
# #                   upper = upper,
# #                   trace=TRUE
# #                   )
# 
#   # An attempt with nlmrt function nlxb. Doesn't provide many good results.
# #   modelCES <- nlxb(model,
# #                    data=dataTable,
# #                    start=list(sigma=sigmaGuess, gamma_E=gamma_EGuess, beta=betaGuess, 
# #                               lambda_L=lambda_LGuess, lambda_E=lambda_EGuess),
# #                    trace=TRUE
# #                   )

@@

    # An unconstrained fit gives rho = -6.33, rho1 = 9.36 (sigma1 = 0.0965) and R^2 = 0.9892.  
    # This result is not economically meaningful.
    # Constraining rho = 0 gives rho1 = -582, again not economically meaningful.
    # So, do a grid search over meaningful regions to see what is best.
    # The result (to with +/- 0.1) is rho = 0.0 and rho1 = 5.7, with R^2 = 0.9885.
    # However, neither the the rho or nor the sigma values are statistically significant.
    rho1 = 5.7
                       rho1 = rho1,
                       rho = rho
  } else if (countryAbbrev == "IR"){
    # Iran works with an unconstrained fit.

@@

  } else if (countryAbbrev == "TZ"){
    # Using the PORT algorithm yields convergence and R^2 = 0.9974 with economically menaingful results.
    # However, none of the rho or sigma values are statistically significant.
                       method="PORT",
  } else if (countryAbbrev == "ZM"){
    # ZM works with the PORT algorithm, but not the LM algorithm.
                       method="PORT",
                       rho=0.0

@@

  return(modelCES)
# print("US U")
# print(summary(cesModel("US", "U")))
# print("UK U")
# print(summary(cesModel("UK", "U")))
# print("JP U")
# print(summary(cesModel("JP", "U")))

# print("")
# print("CN Q")
# print(summary(cesModel("CN", "Q")))
# 
# print("")
# print("ZA Q")
# print(summary(cesModel("ZA", "Q")))
# 
# print("")
# print("SA Q")
# print(summary(cesModel("SA", "Q")))
# 
# print("")
# print("IR Q")
# print(summary(cesModel("IR", "Q")))
# 
# print("")
# print("TZ Q")
# print(summary(cesModel("TZ", "Q")))
# 
# print("")
# print("ZM Q")
# print(summary(cesModel("ZM", "Q")))

#########################
# Takes the CES fitted models and creates per-country predictions for them.
# Returns a data.frame with the predictions.
##
cesPredictions <- function(countryAbbrev, energyType){
  xNames = c("iCapStk", "iLabor", "iEToFit")
  tName = "iYear"
  dataTable <- loadAndPrepDataForCES(countryAbbrev, energyType)
  # Can't make predictions for any of CN, ZA, SA, IR, TZ, or ZM if we're interested in U
  if (!(haveDataSF(countryAbbrev, energyType))){
    # If we don't have data for this combination of countryAbbrev and energyType, 
    # return a column of NAs if the above conditions have been met.
    nRows <- 21 # All of these countries need 21 rows.
    df <- as.data.frame(matrix(NA, ncol = 1, nrow = nRows))
    colnames(df) <- "pred"
    return(df)
  }
  model <- cesModel(countryAbbrev, energyType)
  print(model)
  pred <- cesCalc(xNames=xNames, data=dataTable, coef=coef(model), tName="iYear", nested=TRUE)
  df <- data.frame(pred)
  # Pad with rows as necessary
  df <- padRows(countryAbbrev, df)
  return(df)
}

print(cesPredictions("US", "Q"))
